---
title: "dplyrcompleted"
author: ""
date: "September 22, 2017"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(nycflights13)
library(tidyverse)
library(dplyr)
```

## Data Manipulation

Pick observations by their values (filter()).  

Reorder the rows (arrange()).  

Pick variables by their names (select()).  

Create new variables with functions of existing variables (mutate()).

Collapse many values down to a single summary (summarise()).

## All verbs work similarly:  

The first argument is a data frame.  

The subsequent arguments describe what to do with the data frame, using the variable names (without quotes).  

The result is a new data frame.  

## OK, but what is a tibble?

[Look here for everything you want to know](http://bit.ly/2ffxvfM)  




##Moving on with dplyr.. filter()

filter() allows you to subset observations based on their values.

```{r }

filter(flights, month == 1, day == 1)

```


## R either prints out the results, or saves them to a variable.

If you want to do both, you can wrap the assignment in parentheses:

```{r}
(jan1 <- filter(flights, month == 1, day == 1))

```
##Finding flights that arrived late, but left on time

```{r }

filter( flights, arr_delay >0 , dep_delay <= 0)



```

## arrange()

arrange() works similarly to filter()   
except that instead of selecting rows,   
it changes their order.

```{r}
arrange(flights, year, month, day)
```


## desc()

Use desc() to re-order by a column in descending order:

```{r}
arrange(flights, desc(arr_delay))
```


##select can also be used to make new tables with only the necessary observations 

```{r }


flight.lga <- filter(flights, origin == "LGA") 
select(flight.lga, origin, contains("TIME"))

```


## mutate - add variables

```{r }


flights_sml <- select(flights, origin,
  year:day, ##this selects columns from year to day
  ends_with("delay"), 
  distance, 
  air_time)
```
#now we add columns

```{r }
mutate(flights_sml, origin,
  gain = arr_delay - dep_delay,
  speed = distance / air_time * 60)
#Two columns were added to the dataframe!






```

##transmute - delete columns from the dataframe

```{r }



transmute(flights, origin, 
  gain = arr_delay - dep_delay,
  hours = air_time / 60)



```

## summarize
this gives us the mean delay, even though it's pretty pointless to get it sorted according to carrier 
group_by() and summarize() are used to create useful grouped summaries
```{r }
delays <- flights %>% ##this is the pipe operator. Read it as "then"
 group_by(dest) %>% 
  summarise(
    count = n(),
    dist = mean(distance, na.rm = TRUE),
    delay = mean(arr_delay, na.rm = TRUE)
  ) %>% 
  filter(count > 20, dest != "HNL")
delays

##what this code essentially does is a series of commands using then to link them.
```


##Grouping by multiple variables

```{r }

#By grouping in layers we're peeling off a level of grouping. Look what happens now...
daily <- group_by(flights, year, month, day)

```

##Using the daily dataset...
```{r }
(per_day   <- summarise(daily, flights = n()))


```

##We're reducing the variables...
```{r } 
(per_month <- summarise(per_day, flights = sum(flights)))



```
##We found the total number of flights per year!!

```{r }


(per_year  <- summarise(per_month, flights = sum(flights)))


```





##Won't it be cool to visualize our EDA?

```{r }

## first, save this dataset to rule out NA values. We don't like NA values

not_cancelled <- flights %>% 
  filter(!is.na(dep_delay), !is.na(arr_delay))

delays <- not_cancelled %>% 
  group_by(tailnum) %>% 
  summarise(
    delay = mean(arr_delay)
  )

```


## Let's plot!
```{r }
ggplot(data = delays, mapping = aes(x = delay)) + 
  geom_freqpoly(binwidth = 10)

```



## And from what we see, some flights have an average delay of 300 minutes. These are the kinds of observations we do EDA for. Have fun exploring!

